---
title: "Week3: Plotly example - Total sightings seen during Safari"
author: "Marius Peché"
date: "August 11, 2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(plotly)
```

## Introduction

This is the second Peer Reviewed Assignment for the __Data Science Products__ course in the __John Hobkins Data Science__ specialization on Coursera.


We are asked to create a set of web page slides that incorporates a interactive graph created with the R package _plotly_.


## Data 

Continuing with the theme of the [previous week](https://mpeche.github.io/ProgrammingAssignment9/Week2.html), the data for this project will be the KNP data set.

This data was collected during a five day safari through the __Kruger National Park__ in South Africa. The data contains the following of each sighting:

 * Species Name and Class
 * Time stamp
 * Location information
```{r}
knp <- read.csv("KNP.csv")
species <- read.csv("Species.csv")
knp <- merge(knp,species[,1:4],by="SpeciesUId")
knp <- knp[order(knp$Timestamp),]
knp$date = as.Date(knp$Timestamp)
```

## Plotly introduction

We will create a graph that will display the number of unique species observed during each day of the safari.
The species will be devided into three Classes:

 * Birds
 * Mammals
 * Reptiles
 
```{r warning=FALSE, message=FALSE}
library(plyr)
knp_small <- ddply(knp,c("date","Class"),
             function(x){ length(unique(x$SpeciesUId))})
```

## Creating the table

Here is the code used to create the table:

```{r echo=TRUE, eval=FALSE}
plot_ly(knp_small,
        x=knp_small$date, y=knp_small$V1,
        color=knp_small$Class,
        mode="line")
```

## Final Plotly presentation
```{r warning=FALSE, message=FALSE}
plot_ly(knp_small,
        x=knp_small$date, y=knp_small$V1,
        color=knp_small$Class,
        mode="line")
```
